\section{Conclusion}

This paper presented a hardware-aware framework for real-time 6-DoF object tracking on mobile and embedded devices. The system explores three progressively more capable tracking configurations: a CPU-based algebraic pipeline, a GPU-accelerated feature processing pipeline, and a geometry-aware render-and-track formulation operating on the $\mathrm{SE}(3)$ pose manifold.

Experimental evaluation on mobile devices highlights the trade-offs between these configurations. While the CPU-based pipeline provides an efficient baseline, its performance degrades with increasing image resolution. GPU acceleration significantly improves feature processing throughput, but introduces constraints related to memory transfer overhead and GPU resource utilization. 

To address the limitations of purely algebraic approaches, the proposed geometry-aware pipeline incorporates rendering-based 2D--3D correspondence lifting and pose refinement on $\mathrm{SE}(3)$. In particular, the tension-based formulation introduces adaptive correspondence weighting within a weighted Gauss--Newton framework, improving robustness to noise, outliers, and partial visibility without requiring manually tuned loss functions.

This combination of hardware-aware computation and geometry-aware optimization results in improved pose stability under challenging conditions such as significant viewpoint variation and limited feature visibility. Although the geometry-aware stage introduces additional per-frame computation, it reduces drift and the need for frequent relocalization, leading to more consistent tracking performance over time.

Overall, the results demonstrate that integrating adaptive, geometry-driven optimization with efficient hardware utilization enables robust and scalable real-time tracking on resource-constrained edge devices. This approach provides a practical alternative to both purely algebraic pipelines and computationally intensive SLAM or learning-based methods for mobile augmented reality applications.