\section{Geometry-Aware SE(3) Tracking}

\input{figures/se3}

While algebraic tracking based on homography estimation provides efficient frame-to-frame motion estimation, it lacks explicit geometric constraints and becomes unstable under large viewpoint changes, scale variation, or partial occlusions. To address these limitations, we introduce a geometry-aware render-and-track formulation that explicitly estimates pose on the $\mathrm{SE}(3)$ manifold. As illustrated in Figure~\ref{fig:se3}, the system integrates synthetic rendering with a state machine that dynamically switches between global matching and local optical flow.

\subsection{Multiscale Synthetic Reference Rendering}

Instead of relying solely on 2D templates, the proposed method renders a set of $L$ synthetic reference views of the object at multiple depth (scale) hypotheses. For each level, the system generates:
\begin{itemize}
    \item A textured appearance buffer
    \item A descriptor map for feature matching
    \item A dense coordinate buffer storing object-space points
\end{itemize}

Let $\mathbf{X} \in \mathbb{R}^3$ denote a point on the object surface. During rendering, this point is projected into image space using:
\[
\mathbf{x} \sim \mathbf{K} [\mathbf{R} \mid \mathbf{t}] \mathbf{X}
\]
where $\mathbf{K}$ is the intrinsic matrix and $(\mathbf{R}, \mathbf{t}) \in \mathrm{SE}(3)$.

The coordinate buffer encodes $\mathbf{X}$ per pixel, enabling a direct mapping between image coordinates and 3D object geometry.

\subsection{Dual-Branch State Machine}

To ensure efficiency, incoming frames are processed through a two-branch state machine:

\begin{itemize}
    \item \textbf{Global Matching Branch:} Activated during initialization or recovery. Features are extracted and matched against the multiscale reference stack. The optimal scale is selected as:
    \[
    \ell^* = \arg\max_\ell |\mathcal{M}_\ell|
    \]
    where $\mathcal{M}_\ell$ denotes valid matches at level $\ell$.
    
    \item \textbf{Optical Flow Branch:} During stable tracking, correspondences are propagated using pyramidal Lucas--Kanade (LK) optical flow \cite{lucaskanade1981}. Forward-backward consistency is enforced:
    \[
    \|\mathbf{x}_{t-1} - \hat{\mathbf{x}}_{t-1}\| < \varepsilon
    \]
\end{itemize}

In both branches, correspondences are filtered using RANSAC-based homography validation to retain only geometrically consistent inliers.

\subsection{2D--3D Correspondence Lifting}

Each validated feature correspondence is lifted to a 2D--3D pair:
\[
(\mathbf{x}_i, \mathbf{X}_i)
\]
by querying the coordinate buffer. This eliminates the need for triangulation or back-projection and directly provides constraints for pose estimation.

\subsection{Tension-Based Pose Refinement on $\mathrm{SE}(3)$}

Given correspondences $(\mathbf{x}_i, \mathbf{X}_i)$, we define the reprojection function:
\[
\mathbf{p}_i(\mathbf{T}) = \pi(\mathbf{K}\mathbf{T}\mathbf{X}_i)
\]
and residual:
\[
\mathbf{e}_i = \mathbf{x}_i - \mathbf{p}_i(\mathbf{T}), \quad d_i = \|\mathbf{e}_i\|
\]

\paragraph{Spring Model Formulation}

Each correspondence is modeled as a virtual spring in image space:

\begin{itemize}
    \item Current length: $d_i$
    \item Natural length:
    \[
    L = \min_i d_i
    \]
\end{itemize}

The energy of each spring is defined as:
\[
E_i = \frac{\lambda}{2}(d_i - L)^2
\]

\paragraph{Gradient Derivation}

Applying the chain rule:
\[
\frac{dE_i}{d\boldsymbol{\xi}} =
\frac{dE_i}{dd_i}
\cdot
\frac{dd_i}{d\mathbf{e}_i}
\cdot
\frac{d\mathbf{e}_i}{d\boldsymbol{\xi}}
\]

we obtain:
\[
\frac{dE_i}{dd_i} = \lambda(d_i - L), \quad
\frac{dd_i}{d\mathbf{e}_i} = \frac{\mathbf{e}_i^T}{d_i}, \quad
\frac{d\mathbf{e}_i}{d\boldsymbol{\xi}} = -\mathbf{J}_i
\]

Combining:
\[
\nabla E =
\sum_i \mathbf{J}_i^T
\left(
\lambda \frac{d_i - L}{d_i} \mathbf{e}_i
\right)
\]

\paragraph{Weighted Gauss--Newton Formulation}

Defining the weight:
\[
w_i = \frac{d_i - L}{d_i}
\]

the optimization becomes:
\[
E(\mathbf{T}) = \sum_i w_i \|\mathbf{e}_i\|^2
\]

Linearizing:
\[
\mathbf{e}_i(\boldsymbol{\xi} + \delta)
\approx
\mathbf{e}_i + \mathbf{J}_i \delta
\]

leads to the normal equations:
\[
(\mathbf{J}^T \mathbf{W} \mathbf{J}) \delta =
\mathbf{J}^T \mathbf{W} \mathbf{e}
\]

where $\mathbf{W} = \mathrm{diag}(w_i)$.

\paragraph{Pose Update}

The pose is updated using the exponential map:
\[
\mathbf{T} \leftarrow \exp(\delta^\wedge)\mathbf{T}
\]

\paragraph{Interpretation}

This formulation can be interpreted as a physics-inspired weighted least-squares system:
\begin{itemize}
    \item Small residuals ($d_i \approx L$): low influence
    \item Large residuals: strong corrective force
\end{itemize}

Unlike standard Gauss--Newton or fixed robust losses, the weighting is adaptive and derived directly from the residual structure of the current frame.

\subsection{Failure Handling and System Integration}

The system is designed to be fault-tolerant. If RANSAC produces insufficient inliers or the number of tracked LK points falls below a threshold ($\texttt{LK\_MIN}$), tracking is reset and the system re-enters the global matching branch.

Successful pose estimates increment a confirmation counter, and once a minimum threshold ($\geq \texttt{MIN\_CONFIRM}$) is reached, the system resumes efficient optical flow tracking.

\paragraph{Implementation Details}

The rendering stage is implemented using a lightweight cross-platform abstraction built on the \textit{Sokol} framework. On mobile devices, OpenGL ES~3 is used for efficient GPU execution. This ensures that coordinate buffer generation remains portable while fully exploiting available GPU hardware.