\section{System Overview}

The proposed framework enables efficient and robust 6-DoF tracking on mobile and embedded devices by combining hardware-aware feature processing with geometry-aware pose refinement. The system is structured as a progressive hierarchy of three tracking configurations: (i) a CPU-based algebraic pipeline, (ii) a GPU-accelerated feature processing pipeline, and (iii) a geometry-aware render-and-track framework operating on the $\mathrm{SE}(3)$ pose manifold. Figure~\ref{fig:backbone} illustrates the overall system architecture.

\input{figures/backbone}

All configurations share a common feature-based tracking backbone. Incoming frames are converted to grayscale and processed using a FAST-based keypoint detector, followed by non-maximum suppression. Local binary descriptors are computed for the detected features and matched across frames using efficient Hamming-distance comparisons. The resulting correspondences are used to estimate motion via an algebraic alignment stage based on homography estimation, while pyramidal Lucas--Kanade optical flow provides efficient frame-to-frame tracking of inlier points.

The first configuration executes the entire pipeline on the CPU and leverages SIMD vectorization to accelerate feature extraction, descriptor computation, and matching. The second configuration introduces GPU acceleration by offloading keypoint detection and descriptor construction to programmable shaders, while descriptor matching is performed in parallel using compute shaders. Both configurations retain the same algebraic motion model and primarily differ in their computational efficiency.

The third configuration augments this pipeline with a geometry-aware pose refinement stage. Synthetic reference views of the known object are rendered at multiple pose hypotheses, enabling direct lifting of image correspondences to 2D--3D constraints. These correspondences are used to refine the pose on $\mathrm{SE}(3)$ through an iterative optimization procedure, incorporating adaptive correspondence weighting to improve robustness under noise, outliers, and viewpoint changes.

Together, these configurations provide a unified framework for analyzing the trade-offs between purely algebraic tracking, hardware-accelerated feature processing, and geometry-aware pose refinement on resource-constrained devices.