\begin{abstract}
Real-time 6-DoF pose estimation on mobile and embedded platforms is constrained by both computational limitations and geometric robustness requirements. Dense SLAM pipelines and learned pose regression models typically require substantial compute and memory resources, making them difficult to deploy efficiently on edge devices. Classical feature-based tracking offers a lightweight alternative, but most existing implementations rely on purely algebraic point-matching formulations—such as homography estimation—that lack explicit 3D geometric constraints and often become unstable under viewpoint changes, scale variation, or partial visibility.

In this work, we present VORTEX, a hardware-aware and geometry-aware tracking framework designed specifically for edge devices. We first redesign the foundational algebraic tracking stages through two optimized implementations: a CPU pipeline leveraging SIMD vectorization and a GPU-resident feature pipeline using custom fragment and compute shaders for keypoint detection, descriptor extraction, and parallel descriptor matching. Building on this accelerated feature backbone, we introduce a render-and-track formulation on the SE(3) manifold that incorporates explicit object geometry through synthetic reference rendering and direct 2D–3D correspondence lifting.

To further improve optimization stability, we propose a physics-inspired formulation for SE(3) pose estimation. Instead of directly minimizing reprojection error, each 2D–3D correspondence is modeled as a virtual spring, where the smallest residual defines a natural length and adaptive tension forces guide the optimization. This results in a weighted Gauss–Newton update that emphasizes geometrically consistent correspondences while suppressing outliers and local degeneracies. The formulation provides improved convergence behavior and robustness under partial visibility and non-uniform feature distributions.

Experiments on Android-based mobile hardware demonstrate that while GPU-accelerated algebraic tracking improves computational throughput, the proposed geometry-aware SE(3) refinement with tension-based optimization significantly enhances pose stability and robustness for real-time 6-DoF object tracking without persistent mapping.

\end{abstract}

\keywords{
6-DoF pose estimation \and
Edge-aware vision systems \and
Hardware-aware computer vision \and
GPU-accelerated feature tracking \and
Render-and-track methods \and
SE(3) pose optimization \and
Mobile augmented reality
}